Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-03-11 supplement.tex
Changed to "ka"

In one parameterization of early structure, which we refer to as a “continuous
migration” model, a secondary stem population (stem 2) split from the primary
stem (stem 1) that leads to contemporary humans. Stem 1 contributed to present-day
populations via a series of population splits similar to the single-origin
model, while stem 2 contributed through continuous symmetric migrations with
contemporaneous populations. The symmetrical migration rates could differ
across population pairs and over different epochs. This continuous migration
was allowed until stem 2 disappeared, which occurred as recently as 5kya. We
tested models that both allowed or disallowed migration between stems, i.e.,
before stem 1 split into S/E/W African populations.

@@

As expected, the clean split models consistently provide divergence times $T$
that are more recent than the IM models. Because $T$ and subsequent migration
can be inversely confounded in SFS inference, fixing migration rates to zero
results in more recent inferred $T$ (Fig.~\ref{fig:supp-IM-misspecification}).
Even though each of our four models inferred divergences of all human
populations $\sim$120kya, ancient structure and size changes in the simulations
biases re-inferences and gives inflated estimates of $T$.
Such an effect has been shown to occur when the ancestral
population size increases prior to the population split
\citep{Momigliano2021-th}, which is the case in our inferred single-origin model.
Re-inferred $T$ between the Nama and other populations were $>$200kya under the IM model,
and $\sim$50--130kya under the clean split model.

>>>>

In one parameterization of early structure, which we refer to as a “continuous
migration” model, a secondary stem population (stem 2) split from the primary
stem (stem 1) that leads to contemporary humans. Stem 1 contributed to present-day
populations via a series of population splits similar to the single-origin
model, while stem 2 contributed through continuous symmetric migrations with
contemporaneous populations. The symmetrical migration rates could differ
across population pairs and over different epochs. This continuous migration
was allowed until stem 2 disappeared, which occurred as recently as 5ka. We
tested models that both allowed or disallowed migration between stems, i.e.,
before stem 1 split into S/E/W African populations.

@@

As expected, the clean split models consistently provide divergence times $T$
that are more recent than the IM models. Because $T$ and subsequent migration
can be inversely confounded in SFS inference, fixing migration rates to zero
results in more recent inferred $T$ (Fig.~\ref{fig:supp-IM-misspecification}).
Even though each of our four models inferred divergences of all human
populations $\sim$120ka, ancient structure and size changes in the simulations
biases re-inferences and gives inflated estimates of $T$.
Such an effect has been shown to occur when the ancestral
population size increases prior to the population split
\citep{Momigliano2021-th}, which is the case in our inferred single-origin model.
Re-inferred $T$ between the Nama and other populations were $>$200ka under the IM model,
and $\sim$50--130ka under the clean split model.
